\subsection{Change of State}

\begin{itemize}
    \item \textbf{Gas} has a much \textbf{lower density} than that of the same substance in liquid or solid state.
        \begin{itemize}
            \item Solids and liquids have molecules packed \textbf{in contact} with each other.
            \item Molecules of a gas are on average \textbf{separated by large distances}.
        \end{itemize}
    \item \textbf{Solids} cannot flow.
        \begin{itemize}
            \item Molecules in a solid are \textbf{locked together by strong force bonds}.
            \item Molecules in solids and gas have \textbf{too much kinetic energy} for the force bonds to keep molecules fixed to each other.
        \end{itemize}
\end{itemize}

\subsubsection*{Latent Heat}

\begin{itemize}
    \item A solid \textbf{heated at its melting point}, its atoms vibrate so much they \textbf{break free from each other}.

        The energy needed to melt a solid at its melting point is called the \textbf{latent heat of fusion} - latent heat is \textbf{released} when a liquid solidifies.
    \item A liquid \textbf{heated at its boiling point}, its molecules \textbf{gain enough kinetic energy} to overcome the bonds that hold them close together.

        The energy needed to vaporise a liquid is called \textbf{latent heat of vaporisation}.
\end{itemize}

\textbf{No temperature change} takes place during melting and boiling, even though the solid is being heated.

\textbf{Sublimation} is when solid vaporise directly when heated.

\begin{itemize}
    \item \textbf{Specific latent heat of fusion} $l_f$ of a substance is the energy needed to change the state of unit mass of the substance from solid to liquid without change of temperature.
    \item \textbf{Specific latent heat of vaporisation} of a substance is the energy needed to change the state of unit mass of the substance from liquid to vapour without changing temperature.
\end{itemize}
$$Q=ml$$

The unit of specific latent heat is J\,kg$^{-1}$
