\subsection{Numbering Systems}

\begin{itemize}
    \item An \textbf{integer} is any whole number.\\
        $\mathbb{Z}=\{\dots,-2,-1,0,1,2,\dots\}$
    \item A \textbf{natural number} is a whole number that is used for counting.\\
        $\mathbb{N}=\{0,1,2,\dots\}$
    \item A \textbf{rational number} is any value that can be expressed as a fraction.\\
        $\mathbb{Q}$
    \item An \textbf{irrational number} cannot be expressed as a fraction, and has an endless series of non-repeating digits.

        An irrational number cannot be correctly represented using a finite number of digits, therefore a \textbf{rounding error} will occur.
\end{itemize}

A \textbf{real number} is any natural, rational or irrational number. The \textbf{set of real numbers} $\mathbb{R}$ is defined as the set of all possible real world quantities.

\textbf{Ordinal numbers} describe the numerical position of objects - first, second, etc.
\begin{itemize}
    \item Natural numbers are used for \textbf{counting}.
    \item Real numbers are used for \textbf{measurement}.
    \item Ordinal numbers are used as \textbf{pointers} to a particular element in a sequence, or to define the position of something in a list.
\end{itemize}

\subsubsection*{Number Bases}
\begin{itemize}
    \item \textbf{Denary} uses the digits 0 through 9 and has a base of 10.
    \item \textbf{Binary} uses only the digits 0 and 1 and has a base of 2.
    \item \textbf{Hexadecimal} uses digits 0-9 and letters A to F and has a base of 16.
\end{itemize}

The numbering base can be written as a subscript $11_\text{16}$.

The hexadecimal system is used as the \textbf{shorthand for binary}, since
\begin{itemize}
    \item It is simple to represent a byte in just two digits.
    \item \textbf{Fewer mistakes} are likely to be made in writing a hex number than a string of binary digits.
    \item It is easier for computer users to \textbf{remember a hex number} than a binary number.
\end{itemize}
\textbf{Colour codes} often use hexadecimal to represent the RBG values, as their are easier to remember than a 24-bit binary string.
