\subsection{Speed and Velocity}
\begin{itemize}
    \item \textbf{Displacement} is distance in a given direction.
    \item \textbf{Speed} is defined as change of distance per unit time.
    \item \textbf{Velocity} is defined as change in displacement per unit time, or
        \begin{itemize}
            \item Velocity is speed in a given direction.
        \end{itemize}
\end{itemize}

The unit of speed and velocity is the metre per unit second $\text{ms}^{-1}$.

An object moving at \textbf{constant speed} travels equal distance in equal times.
\begin{itemize}
    \item For an object which travels distance $s$ in time $t$ at constant speed.
        \begin{align*}
            \text{speed}\ v&=\frac{s}{t}\\
            \text{distance travelled}\ s&=vt
        \end{align*}
    \item For an object moving at  constant speed on a circle of radius $r$, its speed
        $$v=\frac{2\pi r}{T}$$
\end{itemize}

For an object moving at \textbf{changing speed} that travels a distance $\Delta s$ in time $\Delta t$
$$v=\frac{\Delta s}{\Delta t}$$
The \textbf{delta notation} $\Delta$ means \underline{a change of} something.

\subsubsection*{Distance-time Graphs}

A distance-time graph is a graph of distance against time.

\begin{itemize}
    \item For an object moving at \textbf{constant speed}, its distance-time graph is a \textbf{straight line with constant gradient}.
    $$\text{speed}\ =\frac{s}{t}=\text{gradient of line}$$
    \item For and object moving at \textbf{changing speed}, the gradient of the line changes.
        \begin{itemize}
            \item The gradient of the line at any point can be found by drawing a \textbf{tangent to the line} at that point.
            \item Then measuring the gradient of the tangent.
        \end{itemize}
\end{itemize}

\subsubsection*{Velocity}

An object moving at \textbf{constant velocity} moves at the same speed without changing its direction of motion.
\begin{itemize}
    \item If an object changes its \textbf{direction of motion} or its \textbf{speed} or both, its velocity changes.
    \item The velocity of an object moving on a \textbf{circular path} at constant speed \textbf{changes continuously} because its direction of motion changes continuously.
\end{itemize}

An object travelling along a straight line has two possible directions, so the \textbf{displacement-time graph} can have a negative gradient when the object moves in the \textbf{negative direction}.
