\subsection{Photons}

\textbf{Visible light} is just a small part of the spectrum of \textbf{electromagnetic waves}. Our eyes cannot detect the other parts.

In a vacuum, all electromagnetic waves travel at the speed of light $c=3.00\times10^8\text{ms}^{-1}$. Where the frequency $f$ and wavelength $\lambda$ of electromagnetic radiation is related by
$$f\lambda=c$$
where the wavelength of light $\lambda$ is often expressed in nanometres ($\text{1nm}=1\times10^{-9}\text{m}$). Visible light have wavelength between 400nm and 700nm.

An electromagnetic wave consists of an \textbf{electric wave} and \textbf{magnetic wave} which travel together and vibrate
\begin{itemize}
    \item At \textbf{right angles to each other} and to the direction which they are travelling.
    \item \textbf{In phase with each other} - the two waves reach a peak together.
\end{itemize}

\subsubsection*{The Photon Theory of Light}

Electromagnetic waves are emitted by a charged particle when \underline{it loses energy}, this happens when
\begin{itemize}
    \item A fast-moving electron is \textbf{stopped}, slowed down or changed direction.
    \item An \textbf{atomic electron} moves to a different shell of lower energy.
\end{itemize}

Electromagnetic waves are emitted in \underline{short bursts of waves}, each burst is a packet of waves referred to as a \textbf{photon}. The \textbf{photon theory} was established to explain the \textbf{photoelectric effect} - the emission of electrons from a metal surface when light is directed at the surface.

The energy $E$ of a photon depends on its frequency $f$
$$E=hf$$
where $h=6.63\times10^{-34}$Js is \textbf{Planck's constant}.

\subsubsection*{Power of Laser Beam}

A \textbf{laser beam} consists of photons of the \underline{same frequency}. The power of a laser beam is the energy per second transferred by the photons.
$$P=nhf$$
where $n$ is the number of photons in the beam passing a fixed point each second.
