\subsection{Particles and Antiparticles}

Matter and antimatter particles destroy each other when they meet, releasing radiation.

In a \textbf{positron emitting tomography} (PET) scanner, a positron-emitting isotope is administered to the patient. Each positron emitted meets an electron and \textbf{annihilate} each other. \underline{Two} gamma photons are produced and sensed by detectors, an image of where the positron-emitting isotopes are is built up from detector signals.

The \textbf{positron} is the antiparticle of the electron.

Positron takes place when a \underline{proton changes into a neutron} in an unstable nucleus with too many protons. The positron $\beta^+$ and an uncharged neutrino $\nu$ are emitted.
$$^A_ZX\to\,^{\hspace{11pt}A}_{Z-1}Y+\,^{\hspace{6pt}0}_{+1}\beta+\nu$$

There are no naturally occurring \textbf{positron-emitting isotopes}. They are manufactured by placing a stable isotope in the path of a beam of protons, some of the nuclei absorb extra protons and become unstable positron-emitters.

\subsubsection*{The Theory of Antiparticles}

$E=mc^2$ shows that the mass of a particle increases the faster it travels. The mass of a particle when it is stationary - its \textbf{rest mass} $m_0$ corresponds to \textbf{rest energy} $m_0c^2$ locked up as mass. It also showed that rest energy must be included in the conservation of energy.

The existence of \textbf{antiparticles} would allow rest energy to be unlocked whenever a particle and a corresponding antiparticle meet and annihilate each other.

The theory of antiparticles predicted that for every type of particle, there is a corresponding antiparticle
\begin{itemize}
    \item \textbf{Annihilate} the particle and itself when they meet, converting their total mass into photons.
    \item Has the \textbf{same rest mass} as the particle.
    \item Has the \textbf{opposite charge} to the particle (if charged).
\end{itemize}

The opposite process of \textbf{pair production} was also predicted - a photon with sufficient energy passing near a nucleus or an electron can suddenly change into a particle-antiparticle pair, which would then be separated from each other.

\subsubsection*{Energy of Annihilation and Pair Production}

The energy of a particle is measured in \textbf{mega electron volts} (MeV). One \textbf{electron volt} (eV) is defined as the energy transferred when an electron is moved through a potential difference of 1V.

\textbf{Annihilation} occurs when a particle and corresponding antiparticle meet
\begin{itemize}
    \item Their mass is converted into \textbf{radiation energy}.
    \item \textbf{Two photons} are produced in the process.
        \begin{itemize}
            \item A single photon cannot ensure a total momentum of zero after the collision.
        \end{itemize}
\end{itemize}

The \textbf{minimum energy} of each photon is given by
$$hf_\text{min}=E_0$$
where $E_0$ is the rest energy of the particle.

In \textbf{pair production}, a photon creates a particle and a corresponding antiparticle, and vanishes in the process. The \textbf{minimum frequency} $f_\text{min}$ of the photon to produce the particle-antiparticle pair is
$$hf_\text{min}=2E_0$$
A proton with less energy cannot create the particle-antiparticle pair.

\subsubsection*{Discovery of the Positron}

A \textbf{cloud chamber} is a small transparent container containing air \textbf{saturated with vapour} and made very cold. \textbf{Ionising particles} leave a visible trail of liquid droplets when they pass through the air, allowing \textbf{cosmic rays} to be photographed.

By allowing particles to pass through a \textbf{lead plate} in the chamber and applying a \textbf{magnetic field}.
\begin{itemize}
    \item A positive particle would be \underline{deflected in the opposite direction} to a negative particle travelling in the same direction.
    \item The slower it went, the more it would bend.
\end{itemize}
where he discovered a beta particle that slowed down but \underline{bent in the opposite direction} to all the other beta trails. He had detected a positron, the first antiparticle to be detected..
