\subsection{Collisions of Electrons with Atoms}

An \textbf{ion} is a charged atom - the number of electrons in an ions is not equal to the number of protons.

An ion if formed from an \textbf{uncharged atom} by adding and removing electrons from the atom. Any process of creating ions is called \textbf{ionisation}
\begin{itemize}
    \item Alpha, beta and gamma radiation create ions when they pass through substances and \textbf{collide with atoms} of the substance.
    \item Electrons passing through a fluorescent tube create ions when they \textbf{collide with atoms} of the vapour in the tube.
\end{itemize}

The \textbf{electron volt} is a unit of energy equal to the work done when an electron is moved through a pd of 1V.

\subsubsection*{Excitation by Collision}

Gas atoms can absorb energy from colliding electrons without being ionised, this process is known as \textbf{excitation}. It happens at \textbf{certain energies}, which are characteristics of the atom.
\begin{itemize}
    \item If a colliding electron \underline{loses all its kinetic energy} when it causes excitation, the current due to the flow of electrons through the gas is \textbf{reduced}.
    \item If a colliding electron \underline{does not have enough kinetic energy} to cause excitation, it is \textbf{deflected by the atom} with no overall loss of kinetic energy.
\end{itemize}

The energy values at which an atom absorbs energy are known as its \textbf{excitation energies}.

The excitations energies of atoms in a gas-filled tube can be determined by
\begin{enumerate}
    \item Increasing the potential difference between the filament and the anode, and
    \item Measure the pd when the anode current falls.
\end{enumerate}

When excitation occurs, the colliding electron makes an atomic electron move from an inner shell to an outer shell. Energy is needed to move the atomic electron away from the nucleus.

The excitation energy is \underline{always less than the ionisation energy} of the atom, because the atomic electron is not removed completely from the atom when excitation occurs.
