\subsection{Specific Heat Capacity}

The temperature rise of an object when heated depends on
\begin{itemize}
    \item \textbf{Mass} of the object.
    \item \textbf{Amount of energy} applied to it.
    \item \textbf{Substance} from which the object is made.
\end{itemize}

The \textbf{specific heat capacity} $c$ of a substance is the energy needed to raise the temperature of unit mass of the substance by 1K without change of state.

The unit of $c$ is J\,kg$^{-1}$\,K$^{-1}$.

The energy needed to raise the temperature of mass $m$ of a substance is
$$\text{Energy needed}\ \Delta Q=mc\Delta t$$

\subsubsection*{Specific Heat Capacity Measurements (Solid)}

\begin{enumerate}
    \item A block of metal of \textbf{known mass} $m$ in an \textbf{insulated container} is used.
    \item An \textbf{electrical heater} is inserted into a hole drilled in the metal.
    \item A \textbf{thermometer} is inserted into a hole drilled in the metal to measure the \textbf{temperature rise}.
    \item A \textbf{small amount of water or oil} in the thermometer hole will \textbf{improve the thermal contact} between the thermometer and the metal.
\end{enumerate}

Assuming \textbf{no heat loss} to the surroundings.
\begin{align*}
    mc\Delta T&=IV\Delta t\\
    c&=\frac{IV\Delta t}{m\Delta T}
\end{align*}

\subsubsection*{Specific Heat Capacity Measurements (Liquid)}

\begin{enumerate}
    \item A \textbf{known mass} of liquid is used in an \textbf{insulated calorimeter} of know mass and known specific heat capacity.
    \item An \textbf{electric heater} is placed in the liquid to heat it directly.
    \item A \textbf{thermometer} is inserted into the liquid to measure the \textbf{temperature rise}.
\end{enumerate}
$$IV\Delta t=m_Lc_L\Delta T+m_\text{cal}c_\text{cal}\Delta T$$
So $c$ can be calculated because all other quantities are known.

\subsubsection*{Continuous Flow Heating}

For mass $m$ of liquid passing through the heater in a time $\Delta t$ at a steady flow rate.
$$IV=\frac{mc\Delta T}{\Delta t}$$
