\subsection{Energy and Simple Harmonic Motion}

A \textbf{freely oscillating object} oscillates with a \textbf{constant amplitude} because there is no friction acting on it - the only forces acting on it combine to provide the \textbf{restoring force}.

\subsubsection*{Velocity of Simple Harmonic Motion}

Consider a small object of mass $m$ oscillating on a spring
\begin{enumerate}
    \item The \textbf{potential energy} changes with displacement $x$ from equilibrium
        $$E_P=\frac{1}{2}kx^2$$
    \item The \textbf{total energy} of the system is equal to its maximum kinetic energy.
        $$E_T=\frac{1}{2}kA^2$$
    \item The total energy is $E_T=E_P+E_K$, so
        $$E_K=E_T-E_P=\frac{1}{2}k(A^2-x^2)$$
    \item Using $E_K=\dfrac{1}{2}mv^2$.
        \begin{align*}
            \frac{1}{2}mv^2&=\frac{1}{2}k(A^2-x^2)\\
            v^2&=\frac{k}{m}(A^2-x^2)\\
            v&=\pm\omega\sqrt{A^2-x^2}\qquad\text{where $\omega=\dfrac{k}{m}$}
        \end{align*}
\end{enumerate}

\subsubsection*{Damped Oscillations}

Oscillations die away because \textbf{air resistance} gradually \textbf{reduces the total energy} of the system.

\begin{itemize}
    \item In any oscillating system where \textbf{friction} or \textbf{air resistance} is present, the amplitude decreases.
    \item These forces are described as \textbf{dissipative forces} because they dissipate energy of the system to the surroundings as thermal energy.
    \item The motion is said to be \textbf{damped} if dissipative forces are present.
\end{itemize}

\textbf{Light damping} occurs when the \textbf{time period is independent of the amplitude} so each cycle takes the same length of time as the oscillations die away.
\begin{itemize}
    \item The displacement of a lightly damped oscillating system \textbf{decreases with time}.
    \item The amplitude \textbf{reduced by the same fraction} each cycle.
\end{itemize}

\textbf{Critical damping} is just enough to stop the system from oscillating after it has been displaced from equilibrium and released.
\begin{itemize}
    \item The oscillating system returns to equilibrium in the \textbf{shortest possible time} without overshooting.
\end{itemize}

\textbf{Heavy damping} is when the damping is so strong the displaced object \textbf{returns to equilibrium more slowly} than if the system is critically damped. \textbf{no oscillating motion} occurs.
