\subsection{The Dangers of Radioactivity}

Ionising radiation is hazardous because it \textbf{damage living cells}.
\begin{itemize}
    \item Destroy cell membrane which causes cells to \textbf{die}.

        High doses of ionising radiation kill living cells.
    \item Damage vital molecules such as DNA to cause cells to \textbf{grow uncontrollably or mutate}.

        Cell mutations and cancerous growth occurs at low and high doses.
\end{itemize}

\subsubsection*{Radiation Monitoring}

Anyone using equipment that produces ionising radiation must wear a \textbf{film badge} to monitor his exposure to ionising radiation.
\begin{itemize}
    \item Different areas of the film are covered by absorbers of different material and different thickness.
    \item When the film is developed, the amount of exposure to each form of ionising radiation can be estimated from the blackening of the film,
    \item If the film is overexposed, the wearer is not allowed to continue working with the equipment.
\end{itemize}

\textbf{Dose} is measured in energy absorbed per unit mass of matter from the radiation.

\subsubsection*{Background Radiation}

Background radiation occurs naturally due to \textbf{cosmic radiation} and from \textbf{radioactive materials}.

\textbf{Storage}: radioactive materials should be in \textbf{lead-lined containers}. The lead lining must be thick enough to reduce the $\gamma$ radiation from the source in the container to about the background level.

No source should be allowed to come into contact with the skin.
\begin{itemize}
    \item \textbf{Solid sources} should be transferred using handling tools or handling robots to ensure the material is \textbf{as far from the user as practicable}.

        So the user is beyond the range of $\alpha$ and $\beta$ radiation of the source, and the $\gamma$ radiation at the user is as low as possible.
    \item Liquid, gas and solid in powdered form should be in \textbf{sealed containers}, so radioactive gas cannot be breathed in, liquid splashed or drunk.
    \item Radioactive sources should not be used for longer than necessary.
\end{itemize}
