\subsection{Planetary Fields}

Newton's law of gravitation can be used to determine the gravitational field strength at any point in the field of \textbf{any spherical mass} - the field of a spherical mass is the same as if the mass were concentrated at its centre.

This is because the field lines of a spherical mass are \textbf{always directed towards the centre}, so the field pattern is just the same as for a point mass.

\begin{itemize}
    \item For a \textbf{point mass $M$}, the magnitude of the force of attraction on a test mass $m$ at distance $r$ from $M$ is given by Newton's law of gravitation.
        $$F=\frac{GMm}{r^2}$$
        The magnitude of the \textbf{gravitational field strength} at distance $r$ is
        $$g=\frac{F}{m}=\frac{GM}{r^2}$$
    \item For a \textbf{spherical mass $M$ and radius $R$}, the force of attraction on a test mass at distance $r$ from $M$ is the same as if mass $M$ is concentrated at its centre.
        $$F=\frac{GMm}{r^2}$$
        So the magnitude of gravitational field strength
        $$g=\frac{F}{m}=\frac{GM}{r^2}$$
        Provided $r\geq R$
\end{itemize}

The \textbf{surface gravitational field strength} $g_s=\dfrac{GM}{R^2}$, so $g=\dfrac{g_sR^2}{r^2}$, for $r\geq R$ the curve is an \textbf{inverse-square law} curve because $g$ decreases in inverse proportion to $r^2$.

For $r<R$, $g\propto r$, more specifically
$$g=\frac{4\pi G\rho r}{3}$$

\subsubsection*{Escape Velocity}

The gravitational potential on Earth's surface is $-63\text{MJ\,kg}^{-1}$, meaning 63MJ of work needs to be done to remove a 1kg mass from the surface of the Earth to infinity.

The \textbf{escape velocity} from a planet is the minimum velocity an object must be given to escape from the planet when projected vertically from the surface.
\begin{enumerate}
    \item The potential at the surface of a planet is $V=-\dfrac{GM}{r}$
    \item To move an object of mass $m$ from the surface to infinity, the work that must be done is $\Delta W=m\Delta V=\dfrac{GMm}{R}$
    \item If the object is projected at speed $v$, for it to be able to escape from the planet, its \textbf{initial kinetic energy}
        $$\frac{1}{2}mv^2\geq m\Delta V$$
\end{enumerate}

Therefore
\begin{align*}
    v^2&\geq\frac{2GM}{R}\\
    v_\text{esc}&=\sqrt{\frac{2GM}{R}}
\end{align*}

Substituting $g=\dfrac{GM}{R^2}$
$$v_\text{esc}=\sqrt{2gR}$$

The work under the curve of $g$ against $r$ represents the work done to move a 1kg mass from infinity to the surface. Therefore the area under the curve gives the value of \textbf{gravitational potential} at the surface.
