\subsection{Circuit Rules}

\subsubsection*{Current Rules}

\textbf{Kirchoff's first rule}: At any junction in a circuit, the total current leaving the junction is equal to the total current entering the junction.

For component \textbf{in series}:

\begin{itemize}
    \item The current entering a component is the \textbf{same as the current leaving} the component.
        \begin{itemize}
            \item Components do not use the current.
        \end{itemize}
    \item The current passing through two or more components in series is the \textbf{same through each component}.
\end{itemize}

\subsubsection*{Potential Difference Rules}

The potential difference between any two points in a circuit is defined as the \textbf{energy transfer per coulomb of charge} that flows from one point to the other.

\begin{itemize}
    \item If the charge carriers lose energy, the potential difference is a \textbf{potential drop}.
    \item Charge carriers gain energy when they \textbf{pass through a cell}, the potential difference is a \textbf{potential rise} equal to the pd across the cell's terminals.
\end{itemize}

\begin{itemize}
    \item For any two components in series, the total pd across all the components is equal to the \textbf{sum of the potential differences} across each component.
        \begin{itemize}
            \item The pd across each component is the \textbf{energy delivered per coulomb charge} to that component.
            \item So the sum of pds across the components is the \textbf{total energy} delivered to the components per coulomb of charge passing through them.
        \end{itemize}
    \item The pd across components in parallel is the same.
\end{itemize}

\textbf{Kirchoff's second law}: For any complete loop of a circuit, the sum of the emfs round the loop is equal to the sum of potential drops around the loop.
\begin{itemize}
    \item The statement follows from the \textbf{conservation of energy}.
    \item The total emf in a loop is the \textbf{total electrical energy} per coulomb produced in the loop.
    \item The sum of the potential drops is the \textbf{electrical energy per coulomb delivered} around the loop.
\end{itemize}
