\subsection{Hubble's Law}

\begin{enumerate}
    \item Hubble measured \textbf{Cepheid variables} which has a known absolute magnitude depending on their period.
    \item He calculated the distance to each Cepheid variable.
    \item He identified that prominent spectral lines are \textbf{red-shifted to longer wavelengths}.
\end{enumerate}

\textbf{Hubble's law} is expressed as
$$v=Hd$$
where $H=65$km\,s$^{-1}$Mpc$^{-1}$ is Hubble's constant.

\subsubsection*{The Big Bang Theory}
\begin{itemize}
    \item \textbf{Big Bang theory} suggest the universe was created in a massive primordial explosion.
    \item \textbf{Steady State theory} supposed that matter enter the Universe at high holes, pushes the galaxies apart as matter enters.
\end{itemize}

\subsubsection*{Evidence for the Big Bang Theory}

\begin{itemize}
    \item \textbf{The spectrum of microwave radiation} from space matched the theoretical spectrum of thermal radiation from an object at a temperature of 2.7K.

        The radiation was created in the Big Bang, travelling through the Universe ever since the Universe became transparent. As the Universe expands, its mean temperature decreases to about 2.7K.
    \item \textbf{Relative abundance of hydrogen and helium} - the 3:1 ratio of hydrogen to helium by mass can be calculated by considering the cooling of the Universe.
\end{itemize}

\subsection*{Dark Energy}

Distant type Ia supernovae are much further than expected.
\begin{itemize}
    \item The supernovae must have been accelerating.
    \item Many more observations confirmed the Universe is accelerating.
    \item The unknown type of force causing this acceleration by releasing hidden energy called \textbf{dark energy}.
\end{itemize}

Evidence for acceleration is based on distance measurements of type Ia supernova by two different methods.
\begin{itemize}
    \item \textbf{Hubble's law} give distance to each supernovae.
    \item Type Ia supernovae used as \textbf{standard candles}.
\end{itemize}

But distant type Ia supernova are dimmer and therefore further away than their red-shifted indicates.
