\subsection{Hardware and Software}

\begin{itemize}
    \item \textbf{Hardware} is the electrical or electro-mechanical parts of a computer and its input/output and storage devices.
    \item \textbf{Software} comprises all the programs that are written to make computers function.
\end{itemize}

Software can be classified into \textbf{system software} and \textbf{application software}.

\subsubsection*{System Software}

System software is the software needed to run the computer's hardware and application programs.
\begin{itemize}
    \item An \textbf{operating system} is a set of programs that lies between application software and the computer hardware.

        Functions include hardware \textbf{resource management} and provision of a \textbf{user interface}.

    \item \textbf{Utility software} \textbf{optimises the performance} of the computer or perform tasks such as backing up files, compressing and encrypting data before transmission, firewall, etc.

        \begin{itemize}
            \item \textbf{Disk defragmenter} reorganises a hard disk so that files which have been split up into blocks and stored all over the disk is \textbf{recombined in a single series of sequential blocks}. This makes reading a file quicker.
            \item \textbf{Virus checker} checks your hard drive and internet downloads for viruses and remove them.
        \end{itemize}

    \item \textbf{Libraries} are ready-compiled programs which can be run when needed.

        Libraries in Windows have a .dll extension.
        
    \item \textbf{Programming language translators} translate program code written by a programmer into machine code which can be run by the computer.
\end{itemize}

\subsubsection*{Application Software}

Application software \textbf{performs specific user-oriented tasks}.
\begin{itemize}
    \item \textbf{General-purpose software} can be used for many different purposes.

        E.g. word-processor, spreadsheet.
    \item \textbf{Special-purpose software} performs a single specific task or set of tasks.

        E.g. hotel booking systems.

    \item Software bought \textbf{off-the-shelf} are ready to use.
    \item \textbf{Bespoke software} package are specifically written by a team of programmers for a particular organisation, to satisfy their particular requirements.
\end{itemize}

Off-the-shelf software
\begin{itemize}
    \item Cheaper to buy, as the cost is shared among all other people buying the package.
    \item It is \textbf{ready to be installed} immediately.
    \item Likely to be \textbf{well-documented, well-tested and bug-free}.
\end{itemize}
It may also contain a lot of features never used.
