\subsection{Transformers}

A transformer changes an alternating pd to a different peak value.
\begin{itemize}
    \item Consists of the \textbf{primary coil} and the \textbf{secondary coil}.
    \item The two coils have the \textbf{same iron coil}.
\end{itemize}

When the primary coil is connected to a source of alternating pd.
\begin{enumerate}
    \item An \textbf{alternating magnetic field} is produced in the core.
    \item The field passes through the secondary coil, so an \textbf{alternating emf} is induced in the secondary coil by the changing magnetic field.
\end{enumerate}

A transformer is designed so that all the magnetic flux produced by the primary coil passes through the secondary coil.

The \textbf{transformer rule}
$$\frac{V_s}{V_p}=\frac{N_s}{N_p}$$
\begin{itemize}
    \item A \textbf{step-up transformer} has $N_s>N_p$ so $V_s>V_p$.
    \item A \textbf{step-down transformer} has $N_s<N_p$ so $V_s<V_p$.
\end{itemize}

\subsubsection*{Transformer Efficiency}

Transformers are almost 100\% efficient because they are designed with
\begin{itemize}
    \item \textbf{Low-resistance} windings to reduce power wasted due to the heating effect of the current.
    \item A \textbf{laminated core} which consists of layers of iron separated by layers of insulator. \textbf{Eddy currents} are reduced so the magnetic flux is as  high as possible - heating effect in the core is reduced.
    \item The \textbf{soft iron core} can be easily magnetised and demagnetised, reduces power wasted through repeated magnetisation and demagnetisation of the core.
\end{itemize}
$$\text{Efficiency}=\frac{I_sV_s}{I_pV_p}$$

\subsubsection*{The Grid System}

The National Grid is a network of transformers and cables which covers all regions of the UK.
\begin{enumerate}
    \item \textbf{Power stations} generates alternating current of 50Hz and 25kV.
    \item \textbf{Step-up transformers} at the power station increases the alternating voltage to 400kV for long-distance transmission via the grid.
    \item Homes are supplied via a \textbf{local transformer substation} at 230V.
\end{enumerate}

Transmission of electrical power over long distances is much more efficient at high voltages than at low voltages.
